---
title: "Homework 2 SDS315"
author: "Jade Lightfoot"
output:
  html_document:
    toc: yes
    toc_float: yes
---
```{r echo=FALSE}
knitr::opts_chunk$set(fig.height=6, fig.width=6, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))

options(dplyr.summarise.inform = FALSE)

```

```{r echo=FALSE , message=FALSE}
# Necessary packages for code 
library(tidyverse)
library(tidyr)
library(ggplot2)
library(dplyr)
library(kableExtra)
library(knitr)
profs<- read_csv("/Users/jadelightfoot/Downloads/profs.csv")
bikeshare <- read_csv("/Users/jadelightfoot/Downloads/bikeshare.csv")
capmetro_UT <- read_csv("/Users/jadelightfoot/Downloads/capmetro_UT.csv")
billboard <- read_csv("/Users/jadelightfoot/Downloads/billboard.csv")
```



## **Problem 1: Beauty, or not, in the classroom** ##

```{r echo=FALSE , eval=FALSE}
# Looking at statistics for graph below
evalscoretab <-table(profs$eval)
maxscore <- names(evalscoretab)[which.max(evalscoretab)]
print(maxscore)
mean(profs$eval)
median(profs$eval)
```

***

### **Part A** ###

```{r echo=FALSE}
# Course Evaluation Scores histograms, frequency of evaluation scores at UT Austin
ggplot(profs, aes(x= eval)) + geom_histogram(binwidth = .1 , fill= "darkorange3", color= "darkorange4") + labs(title = "Course Evaluation Scores (UT Austin)" , x = "Evaluation Score", y= "Count")

```

The graph above displays a histogram of course evaluation scores at The University of Texas at Austin. The range at which students could score courses scales from 1 to 5. Based on this plot most students evaluated courses at the university a 4 and it was rare for students to score courses form 1 to 2. 

***
### **Part B** ###


```{r echo=FALSE}
# Box plot of Course Evaluation Scores by Native Status 
boxplot(profs$eval ~ profs$native, 
        main = "Boxplot of Course Evaluation Scores by Native Status",  xlab = "Native Status", ylab = "Score", names = c("Not Native", "Native"), col = c("sandybrown", "salmon1"))



```

This graph displays the distribution among course evaluation scores based upon whether or no the professor is or is not a native English speaker. When observing this plot it suggest that native English speakers generally had a larger range and higher overall course evaluation scores when compared to non native English speakers. 


***
### **Part C** ###

```{r echo=FALSE}
# Distibution of Course Evaluation Scores by gender 
ggplot(profs, aes(x=eval)) + geom_histogram(binwidth = .5, fill= "salmon2", color="salmon4") + facet_wrap(~gender, ncol=2) + labs(title= "Distribution of Course Evaluation Scores by Gender", x="Score", y="Count") 


```

The histograms show the distribution of course evaluation scores by gender (male or female) among professors. Based on the two different plots students evaluations scores for male professors consisted of a larger number of scores that were a 5. Overall male professors had a higher count of evaluation scores but the histograms for both male and female maintained the same general shape.  


***
### **Part D** ###
```{r echo=FALSE , eval=FALSE}
# Correlation between physical attractiveness and course evaluation score
cor(profs$beauty, profs$eval)

```


```{r echo=FALSE}
# Association between Physical Attractiveness and Course Evaluation Score
ggplot(profs, aes(x=beauty, y=eval)) + geom_point(color= "coral3") + labs(title = "Association Between Physical Attractiveness and Course Evaluations", x = "Physical Attractiveness", y = "Course Evaluation Score")

```

The relationship between course evaluation scores and physical attractiveness of a professor has no association between each other. Meaning that the attractiveness of a professor has little to nothing to do with how the course they teach is scored. Due to the random distribution of points on the graph there is no association. 


## **Problem 2: Bike Sharing** ##

***

### **Plot A** ###

```{r echo=FALSE}
# Grouping bike rentals by hour in the day and calculating the mean of the total variable in each group. 

avg_hr_rentals <- bikeshare %>%
  group_by(hr) %>%
  summarise(avg_total=mean(total)) 
```

```{r echo= FALSE}
#Line graph for Average Hourly Bike Rentals 
ggplot(avg_hr_rentals, aes(x=hr , y=avg_total)) + geom_line(color = "cadetblue") + labs(title= "Average Hourly Bike Rentals", x= "Hours of the day (hr)", y= "Average Total Bike Rentals")
```

This line graph shows the average hourly bike rentals in Washington DC from 2011 and 2012 across all hours of the day. The x- axis explains the hours of the day with 0 being 12:00 am. The y- axis represents the total average number of bikes rented. When looking at the graph one can observe the two peaks, which could be a result from high volume traffic in the morning as people travel to work or school, and in the evening when people are leaving work or school. This observation can be made due to the fact that around the hours 6 to 8 am there is an obvious peak then followed by a drop mid day and then picks back up around the time around 6 to 8 pm. This general pattern follows the concentration of people and high traffic travel times.

***

### **Plot B** ###

```{r echo=FALSE , warning=FALSE , message=FALSE}
#Grouping bike rentals by hour and working day 
avgrentworkday <- bikeshare %>%
  group_by(hr, workingday) %>%
  summarise(avg_total = mean(total))
```

```{r echo=FALSE}
# Faceted Graphs for working days-- Average Hourly Bike Rentals
ggplot(avgrentworkday, aes(x=hr, y=avg_total)) +
 geom_line(color = "dodgerblue") + facet_wrap(~workingday) + labs(title="Average Hourly Bike Rentals", x="Hour of the Day", y="Average Total Bike Rentals")
```

The graphs above display average total bike rentals by hour of the day. The graphs are split based on working day; if the day is neither weekend or holiday it is classified as 1 and 0 otherwise. When looking at the graph on the left bike rentals remain generally low until later on in the day bike rentals peak. The factors that explain this peak on the weekends or holidays or specifically unknown, but is there are lower numbers of bikes rented during this time. Bike rentals could increase at its peak in graph "0" because of mid day hours, where people and families are usually going out. While the graph on the right, graph "1", follows the same pattern pictured in Plot A. There are two distinct peaks in plot "1", which represents the weekdays excluding holidays, and weekends. The peaks could be due to high traffic travel times for people as they commute to their jobs, school, etc. 

***

### **Plot C** ###

```{r echo=FALSE , messages=FALSE , warning=FALSE}
# Grouping the average total bike rentals for 9am 
nineam_data <- filter(bikeshare, hr== 9)
average_ridership <- nineam_data %>%
  group_by(weathersit, workingday) %>%
  summarise(avg_total = mean(total))
```

```{r echo= FALSE}
ggplot(average_ridership, aes(x=weathersit, y=avg_total)) +
 geom_bar(stat='identity', color="paleturquoise3", fill= "paleturquoise3") + facet_wrap(~workingday) + labs(title="Average Ridership During 9 AM", x="Weather Situation", y="Average Total Bike Rentals")
```

The bar plot showcases average ridership during 9am by weather situation and whether it is a working day or not. The graphs are split in two based on working day; if the day is neither weekend or holiday it is classified as 1 and 0 otherwise. The x- axis represents the weather situation during the 9 am hour. "1" representing Clear, Few clouds, Partly cloudy, Partly cloudy, "2" Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist, "3" Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds, and "4" Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog. The y- axis represents average total bike rentals during the 9am hour and certain weather conditions. Based on these bar graphs one can notice that when it is snowing, light rain, thunderstorms, or fog the number of bike rentals decreases for both graphs. 

## **Problem 3: Capital Metro UT Ridership** ##

***

### **Faceted Line Graph** ###
```{r echo= FALSE}
# Recode the categorical variables in sensible, rather than alphabetical, order
capmetro_UT = mutate(capmetro_UT,
day_of_week = factor(day_of_week,
levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
month = factor(month,
levels=c("Sep", "Oct","Nov")))

```

```{r echo= FALSE}
cap_avg <- capmetro_UT %>%
 group_by(hour_of_day, day_of_week, month) %>%
 summarise(avg_boarding = mean(boarding, na.rm = TRUE))
```

```{r echo=FALSE}
ggplot(cap_avg, aes(x = hour_of_day, y = avg_boarding, color = month)) + geom_line() + facet_wrap(~day_of_week) + labs(x= "Hour of Day", y="Average Boardings")

```
The figure shown displays information on average boardings on Austin's Capital Metro bus network from September to November by day of the week and hour of the day. The hour of peak boardings do not change from day to day during the weekdays, but during the weekends (Saturday, Sunday) there is no obvious peak for all the months in average boardings. The average boardings on Monday in September may look lower, could be due to labor day, in which during holiday times like labor day people are less likely to board the Cap Metro. Also the possible reason that average boardings on Wednesday, Thursday, and Friday in November look lower is possibly due to the start of Thanksgiving break for students around UT and the holidays for others. 

***

### **Faceted Scatter Plot** ###

```{r echo=FALSE}
ggplot(capmetro_UT, aes(x = temperature, y = boarding, colour = weekend)) +
 geom_point() +
 facet_wrap(~hour_of_day)

```

In the figure above it shows multiple scatterplots displaying boardings versus temperature which are plotted by hour of the day starting at 6am to 9pm. The dots are red for the weekday and blue for the weekend. Based on all the plots, the hour of the day and temperature outside has no effect on the number of UT students riding the bus. 

## **Problem 4: Wrangling the Billboard Top 100** ##

***

### **Part A** ###

```{r echo= FALSE}
billboard_weeks <- billboard %>%
 group_by(performer, song) %>%
 summarise(count = sum(week_position, na.rm = TRUE))

billboard_sorted <- billboard_weeks %>%
 arrange(desc(count))
```

```{r echo= FALSE }
top_ten <- head(billboard_sorted, 10)
```

```{r echo=FALSE}
kable(top_ten, "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = F) 
```
In the table above it displays the 10 most popular songs since 1958 by the total number of weeks that a song spent on the Billboard Top 100. 

***

### **Part B** ###

```{r echo=FALSE}
bill_filtered <- billboard %>% 
  filter(!(year %in% c(1958, 2021)))


diversity <- bill_filtered %>% 
 group_by(year) %>% 
 summarise(unique_songs = length(unique(song)))
```

```{r echo=FALSE}
ggplot(diversity, aes(x = year, y = unique_songs)) +geom_line() + labs(x = "Year", y = "Number of Unique Songs", title = "Musical Diversity Over Time")
```

The line graph above showcases music diversity over time (1958-2021). Interestingly, there is a huge dip in music diversity in the year 2000, which could be caused in a boom of a specific song. The uniqueness of a song is identified by how many times it appeared on the Top 100.

***

### **Part C** ###

```{r echo=FALSE}  
billtenweek <- billboard %>%
  filter(weeks_on_chart == 10) %>%
  group_by(performer) %>%
  summarize(count= length(performer)) %>%
  filter(count >= 30) %>%
  arrange(desc(count))


ggplot(billtenweek) + geom_col(aes(x= performer, y= count)) + coord_flip() + labs(title= "Artist that Have Ten-Week Hits", x= "Performer", y= "Count")

```

The bar plot above showcases musical artist that have 30 or more songs that have been in the Billboard Top 100 for 10 weeks. 





